Counting And Confusion: Bayesian Rate Estimation With Multiple Populations

### 2013-01-17 many.tex
Adding small changes based on September e-mail exchange.
Minor acknowledgements changes and address changes prior to P&P submission.

\author{Curt Cutler} \email{Curt.J.Cutler@jpl.nasa.gov}
\affiliation{Jet Propulsion Lab, 4800 Oak Grove Dr., Pasadena, CA
  91109} \affiliation{Theoretical Astrophysics, California Institute
  of Technology, Pasadena, California 91125}

@@

If $\tilde{b}(x_N,\theta) \ll 1$, we obtain the result $(1-
\hat{F}(x_N,\theta))\,R_f \approx 1/2$. This can be understood as the
statement that the rate of foreground events with ranking statistic
greater than $x_N$, $(1- \hat{F}(x_N,\theta))\,R_f$, is of order $1$,
as expected. However, $\tilde{b}(x_N,\theta) = -{\rm d}
[\ln(1-\hat{B}(x,\theta))]/{\rm d}x$ and $(1-\hat{B}(x,\theta))
\rightarrow 0$ as $x \rightarrow \infty$, so this term will be
divergent and for many reasonable examples, we will find
$\tilde{b}(x_N,\theta) >> \hat{f}(x_N,\theta)$, in which case the
posterior on $R_f$ is peaked at $0$. This issue highlights the problem
with using a loudest-event statistic with an improper prior on the
background rate $R_b$. No matter how improbable an event with $x=x_N$
is under the background distribution, it can become likely that the
event at $x_N$ is from the background distribution by taking the
background rate to be sufficiently large. Although this predicts many
more events with $x < x_N$, by using only the loudest event we do not
incorporate the information that no such events are seen. This problem
is avoided in the new framework described here, since we use all
events detected above threshold and combined rates, $R_f+R_b$,
significantly greater than the total number of observed events are
strongly disfavored.

@@

\begin{acknowledgments}
  We thank Kipp Cannon, Chad Hanna, Drew Keppel, Richard
  O'Shaughnessy, and Ted von Hippel for discussions and suggestions
  about this manuscript.  IM and WMF acknowledge the hospitality of KITP, supported in part by the National Science Foundation under No. NSF Grant PHY11-25915.  CC's work was carried out  at the Jet Propulsion Laboratory, California Institute of Technology, under contract to the National Aeronautics and Space Administration.  CC also gratefully acknowledges support from NSF grant PHY1068881.  JG's work is supported by the Royal Society.  Copyright 2013.  All rights reserved.
\end{acknowledgments}

>>>>

\author{Curt Cutler} \email{Curt.J.Cutler@jpl.nasa.gov}
\affiliation{Jet Propulsion Lab \\ 4800 Oak Grove Dr \\ Pasadena, CA
  91109} \affiliation{Theoretical Astrophysics \\ California Institute
  of Technology \\ Pasadena, CA 91125}

@@

If $\tilde{b}(x_N,\theta) \ll \tilde{f}(x_N,\theta)$, we obtain the result $(1-
\hat{F}(x_N,\theta))\,R_f \approx 1/2$. This can be understood as the
statement that the rate of foreground events with ranking statistic
greater than $x_N$, $(1- \hat{F}(x_N,\theta))\,R_f$, is of order $1$,
as expected. However, $\tilde{b}(x_N,\theta) = -{\rm d}
[\ln(1-\hat{B}(x,\theta))]/{\rm d}x$ and $(1-\hat{B}(x,\theta))
\rightarrow 0$ as $x \rightarrow \infty$, so this term may be
divergent and for many reasonable examples, we will find
$\tilde{b}(x_N,\theta) \gg \tilde{f}(x_N,\theta)$, in which case the
posterior on $R_f$ is peaked at $0$. This issue highlights the problem
with using a loudest-event statistic with an improper prior on the
background rate $R_b$. No matter how improbable an event with $x=x_N$
is under the background distribution, it can become likely that the
event at $x_N$ is from the background distribution by taking the
background rate to be sufficiently large. Although this predicts many
more events with $x < x_N$, by using only the loudest event we do not
incorporate the information that no such events are seen. This problem
is avoided in the new framework described here, since we use all
events detected above threshold and combined rates, $R_f+R_b$,
significantly greater than the total number of observed events are
strongly disfavored.

@@

\begin{acknowledgments}
  We thank Kipp Cannon, Chad Hanna, Drew Keppel, Richard
  O'Shaughnessy, and Ted von Hippel for discussions and suggestions
  about this manuscript.  IM and WMF acknowledge the hospitality of
  KITP, supported in part by the National Science Foundation under NSF
  Grant PHY11-25915.  CC's work was carried out at the Jet Propulsion
  Laboratory, California Institute of Technology, under contract to
  the National Aeronautics and Space Administration.  CC also
  gratefully acknowledges support from NSF grant PHY1068881.  JG's
  work is supported by the Royal Society.  Copyright 2013.  All rights
  reserved.
\end{acknowledgments}
